---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "inst/README-"
)
```

kdecopula
=========
> Kernel smoothing for bivariate copula densities

[![Build status Linux](https://travis-ci.org/tnagler/kdecopula.svg?branch=master)](https://travis-ci.org/tnagler/kdecopula)
[![Windows Build status](http://ci.appveyor.com/api/projects/status/github/tnagler/kdecopula?svg=true)](https://ci.appveyor.com/project/tnagler/kdecopula)
[![CRAN version](http://www.r-pkg.org/badges/version/kdecopula)](https://cran.r-project.org/web/packages/kdecopula/index.html)
[![CRAN downloads](http://cranlogs.r-pkg.org/badges/kdecopula)](https://cran.r-project.org/web/packages/kdecopula/index.html)  

This package provides fast implementations of kernel estimators for the copula 
density. Due to its several plotting options it is particularly useful for the 
exploratory analysis of dependece structures. It can be further used for flexible
nonparametric estimation of copula densities and resampling.

You can install:

* the stable release on CRAN:
``` r
install.pacakges("kdecopula")
````

* the latest development version:
``` r
devtools::install_github("tnagler/kdecopula")
````


----------------------------------------------------------------

## Functions

The package provides the following functions:

  * `kdecop`: Kernel estimation of a copula density. By default, estimation
  method and bandwidth are selected automatically. Returns an object of class
  `kdecopula`.
  
  * `dkdecop`: Evaluates the density of a `kdecopula` object.
  
  * `pkdecop`: Evaluates the distribution function of a `kdecopula` object.
  
  * `rkdecop`: Simulates synthetic data from a `kdecopula` object.
  
  * Methods for class `kdecopula`: 
  
    * `plot`, `contour`: Surface and contour plots of the density estimate. 
    
    * `print`, `summary`: Displays further information about the density
      estimate.    
      
    * `logLik`, `AIC`, `BIC`: Extracts fit statistics.

Please look up the package documentation for more details on arguments and
options. 


----------------------------------------------------------------

## kdecopula in action

```{r echo = FALSE}
set.seed(1)
```

Below, we demonstrate the main capabilities of the `kdecopula` 
package. All user-level functions will be introduced with small examples. 

Let's consider some variables of the *Wiscon diagnostic breast cancer* data 
included in this package. The data are transformed to pseudo-observations of
the copula by the empirical probability/rank transform:


```{r fig.width=4, fig.height=4}
library(kdecopula)
data(wdbc)  # load data
u <- apply(wdbc[, c(2, 8)], 2, rank) / (nrow(wdbc) + 1)  # empirical PIT
plot(u)  # scatter plot
```

We see that the data are slightly asymmetric w.r.t. both diagonals. Common
parametric copula models are usually not flexible enough to reflect this. Let's
see how a kernel estimator does.


#### Estimation of bivariate copula densities

We start by estimating the copula density with the `kdecop` function. There is
a number of options for the smoothing parameterization, estimation method and 
evaluation grid, but it is only required to provide a data-matrix.

```{r}
kde.fit <- kdecop(u)  # kernel estimation (bandwidth selected automatically)
summary(kde.fit)
```

The output of the function `kdecop` is an object of class `kdecopula` that 
contains all information collected during the estimation process and summary
statistics such as *AIC* or the *effective number of parameters/degrees of
freedom*. These can also be accessed directly, e.g.
```{r}
logLik(kde.fit)
AIC(kde.fit)
```

#### Plotting bivariate copula densities

The most interesting part for most people is probably to make exploratory plots. 
The class `kdecopula` has its own generic for plotting. In general, there are
two possible types of plots: *contour* and *surface* (or perspective) plots. 
Additionally, the `margins` argument allows to choose between a plot of the 
original copula density and a meta-copula density with standard normal margins 
(default for `type = surface`).

```{r fig.width=4, fig.height=4}
plot(kde.fit)
```
```{r fig.width=4, fig.height=4}
contour(kde.fit)
```
```{r fig.width=4, fig.height=4}
contour(kde.fit, margins = "unif")
```

You can also pass further arguments to the `...` argument to refine the
aesthetics. The arguments are forwaded to  
`lattice::wireframe` or `graphics::contour`, respectively.
```{r fig.width=4, fig.height=4}
plot(kde.fit, 
     size = 15,  # reduce grid size
     zlim = c(0, 10),  # z-axis limits
     screen = list(x = -75, z = 45),  # rotate screen
     xlab = list(quote(u[1]), rot =  25),  # labels have to be rotated as well
     ylab = list(quote(u[2]), rot = -25))  
```
```{r fig.width=4, fig.height=4}
contour(kde.fit, col = terrain.colors(30), levels = seq(0, 0.3, by = 0.01))
```

We see that the asymmetries observed in the data are adequately reflected by
the estimated model.


#### Working with a `kdecopula` object

The density and *cdf* can be computed easily:

```{r}
dkdecop(c(0.1, 0.2), kde.fit)  # estimated copula density
pkdecop(cbind(c(0.1, 0.9), c(0.1, 0.9)), kde.fit) # corresponding copula cdf
```

Furthermore, we can simulate synthetic data from the estimated density:
```{r fig.width=4, fig.height=4}
unew <- rkdecop(655, kde.fit)
plot(unew)
```

## References

Gijbels, I. and Mielniczuk, J. (1990). Estimating the density of a copula 
function. *Communications in Statistics - Theory and Methods*, 19(2):445-464. 

Charpentier, A., Fermanian, J.-D., and Scaillet, O. (2006). The estimation of 
copulas: Theory and practice. In Rank, J., editor, Copulas: From theory to
application in finance. Risk Books. 

Geenens, G., Charpentier, A., and Paindaveine, D. (2014). Probit transformation
for nonparametric kernel estimation of the copula density. 
[*arXiv:1404.4414 (stat.ME)*](arxiv.org/abs/1404.4414).

Nagler, T. (2014). Kernel Methods for Vine Copula Estimation. [*Master's Thesis,
Technische Universität München*](https://mediatum.ub.tum.de/node?id=1231221)


